---
title: "Kriging analysis"
author: "Seonga Cho"
date: "12/3/2020"
output: html_document
---


#Install and read in packages
```{r}
library(spdep)
library(spatialreg)
require(spdplyr) || install.packages("spdplyr", dependencies = T)
library(spdplyr)
require(rgdal) || install.packages("rgdal", dependencies = T)
library(rgdal)
require(maptools) || install.packages("maptools", dependencies = T)
library(maptools)
require(dplyr) || install.packages("dplyr", dependencies = T)
library(dplyr)
require(igraph) || install.packages("igraph", dependencies = T)
library(igraph)
require(foreign) || install.packages("foreign", dependencies = T)
library(foreign)
require(maptools) || install.packages("maptools", dependencies = T)
library(maptools)
require(rgeos) || install.packages("rgeos", dependencies = T)
library(rgeos)
require(sp) || install.packages("sp", dependencies = T)
library(sp)
require(ggplot2) || install.packages("ggplot2", dependencies = T)
library(ggplot2)
library(raster)
library(lubridate)
require(sf) || install.packages("sf", dependencies = T)
library(sf)
require(clue) || install.packages("clue", dependencies = T)
library(clue)
require(ggmap) || install.packages("ggmap", dependencies = T)
library(ggmap)
require(plotly) || install.packages("plotly", dependencies = T)
library(plotly)
library(spdep)
options(scipen = 999)
```

#Make sure necessary packages are loaded
```{r}
library(sp)
library(gstat)
```

#read in data
```{r}
metadata <- read.csv("~/PSTAT236/PSTAT236_metadata.csv", stringsAsFactors = F)
metadata <- metadata [-c (22, 23)]
```

#Check structure of data and do initial clean up
```{r}
str(metadata)

# convert continuous variables to numeric that were read in as characters
metadata$amphibian_svl <- as.numeric (metadata$amphibian_svl)
metadata$amphibian_weight <- as.numeric (metadata$amphibian_weight)

# remove outlier
metadata <- metadata[-201, ]

# convert qpcr loads to zoospore equivalents 
metadata$avg_inf <- metadata$average_qpcr * 80 
```

#Create interpolation grid
```{r}
library(rgdal)

# read in shapefiles                               
grid <- readOGR("~/PSTAT236/grid_meta_26910.shp")
frog <- readOGR("~/PSTAT236/meta_data_26910.shp") 
                               
grid_coord <- as.data.frame(grid@coords)
                               
gridded (grid_coord) = ~coords.x1 + coords.x2
                               
frog@data$qpcr_a = as.numeric(frog@data$qpcr_a)
frog@data$qpcr_b = as.numeric(frog@data$qpcr_b)
frog@data$qpcr_c = as.numeric(frog@data$qpcr_c)
                               
                               
frog@data$amphibian_svl <- as.numeric(frog@data$amphibia_2)
metadata$amphibian_weight <- as.numeric (frog@data$amphibia_3)
                               
frog@data$avg_inf <- as.numeric(frog@data$average_qp) * 80 
                               
frog_not_na = frog[which(!is.na(frog@data$avg_inf)),]
frog_pos = frog[which(frog@data$avg_inf >= 1.0),]
                               
frog_pos2 = frog_pos[which(frog_pos@data$avg_inf < 1000),]
```                               

#Plot histograms of target variable (infection load) and log10 transformation
```{r}
# removing NAs from dataset
df_no_na <- metadata[!is.na(metadata$avg_inf), ]

# now I'm subsetting the dataframe we just made to include only positive infections
df_pos <- df_no_na[df_no_na$avg_inf >= 1.0, ]

# remove rows where lat/long are missing
df_pos <- df_pos[!is.na(df_pos$lat), ]

require(lattice)

h_1 <- histogram (~ avg_inf, df_pos, xlab = "Infection load", col = "thistle3")
h_2 <- histogram (~ log10(avg_inf), df_pos, xlab = "log10(infection load)", col = "thistle3")

print (h_1, split = c(1, 1, 2, 1), more = T)
print (h_2, split = c(2, 1, 2, 1), more = F)

# remove these variables
rm (h_1, h_2)
```
We can see from these histograms that the untransformed infection loads are highly skewed to the left, due to the large proportion of low infection loads. However, log10-transforming the data results in a more normal distribution, so we will move forward working with the log10 transformed infection loads.

# look at the proportions higher than each level
```{r pressure, echo=TRUE}
# show the proportions higher than the various thresholds
ph <- function (level) {
  round (100 * sum (df_pos$avg_inf > level) / length (df_pos$avg_inf), 1)
}

p <- NULL; lvls <- c(100, 1000, 10000, 100000, 1000000)

for (l in lvls) p <- c(p, ph(l))

# display table of thresholds and proportions
(data.frame (cbind (level = lvls, percent.higher = p)))

rm (ph, l, lvls, p)
```


# Covariables
```{r pressure, echo=TRUE}
frog_pos@data$amphibia_1 = as.numeric(frog_pos@data$amphibia_1)
frog_pos@data$amphibia_2 = as.numeric(frog_pos@data$amphibia_2)
frog_pos@data$ph = as.numeric(frog_pos@data$ph)
frog_pos@data$watertemp = as.numeric(frog_pos@data$watertemp)
frog_pos@data$conductivi = as.numeric(frog_pos@data$conductivi)
frog_pos@data$totaldisso = as.numeric(frog_pos@data$totaldisso)
frog_pos@data$salinity_p = as.numeric(frog_pos@data$salinity_p)
```

Plot autogenerated variogram
```{r}
library(automap)

# plot variogram to estimate variogram model form and parameters                
v1 = autofitVariogram(log10(avg_inf) ~ 1, frog_pos)
plot(v1)

m <- vgm(1.2, "Sph", 452, 1) 
grid_coord@proj4string = frog_pos@proj4string

# this is a small sample and the variogram does not seem to have great structure
```

#Plot histograms of co-variables
```{r}
h1 <- histogram (~ amphibian_svl, df_pos, xlab = "SVL", col = "#d53e4f")
h2 <- histogram (~ amphibian_weight, df_pos, xlab = "Mass", col = "#fee08b")
h3 <- histogram (~ ph, df_pos, xlab = "Ph", col = "#99d594")
h4 <- histogram (~ watertemp, df_pos, xlab = "Water temperature", col = "#3288bd")
h5 <- histogram (~ conductivity_uS, df_pos, xlab = "Water conductivity", col = "#998ec3")
h6 <- histogram (~ totaldissolvedsolid, df_pos, xlab = "Total dissolved solids", col = "#8c510a")
h7 <- histogram (~ salinity_ppt, df_pos, xlab = "Salinity", col = "#fc8d59")

print (h1, split = c(1, 1, 4, 2), more = T)
print (h2, split = c(2, 1, 4, 2), more = T)
print (h3, split = c(3, 1, 4, 2), more = T)
print (h4, split = c(4, 1, 4, 2), more = T)
print (h5, split = c(1, 2, 4, 2), more = T)
print (h6, split = c(2, 2, 4, 2), more = T)
print (h7, split = c(3, 2, 4, 2), more = F)

rm (h1, h2, h3, h4, h5, h6)
```

#Plot variograms
```{r pressure, echo=TRUE} 
# log(10) infection load variograms
ze_log = variogram(log10(avg_inf) ~ 1, frog_pos)

# plot variogram cloud 
plot(ze_log)
# this doesn't look like there is any spatial correlation within infection loads

## Try different variogram models to see if they fit any better
# Spherical Model
v.sph = fit.variogram(ze_log, vgm(1, "Sph", 30000, 1)) 

# Exponential Model
v.exp = fit.variogram(ze_log, vgm(10, "Exp", 30000, 1)) 

# gaussian model
v.gau = fit.variogram(ze_log, vgm(4, "Gau", 30000, 1)) 

# plot these variograms                       
plot(ze_log, v.sph)
plot(ze_log, v.exp)
plot(ze_log, v.gau)

#### best fit is with the exponential model

# untransformed infection load variograms                          
ze = variogram(avg_inf ~ 1, frog_pos)
plot(ze)

# Spherical Model
v.sph = fit.variogram(ze, vgm(10000000000, "Sph", 30000, 1)) 

# Exponential Model
v.exp = fit.variogram(ze, vgm(10000000000, "Exp", 30000, 1))

# gaussian model
v.gau = fit.variogram(ze, vgm(10000000000, "Gau", 30000, 1)) 

# plot variograms                  
plot(ze, v.sph)
plot(ze, v.exp)
plot(ze, v.gau)
```

# start kriging analyses
```{r}
## Ordinary kriging

### Use fitted model of regionalization to interpolate log(10) infection load with ordinary kriging on the prediction grid.

# interpolate
Ordinary.K <- krige(log10(avg_inf) ~ 1, frog_pos, grid_coord, model = m)
Ordinary.k.d. <- data.frame(Ordinary.K)

# summary statistics
summary (k.o)
```

### Display map of predictions and prediction errors
```{r}
# this file for producing these plots should be found on Rossiter's website, but we could not find the file there or anywhere else online. Therefore, we were unable to run this part
source ("ck_plotfns.R")
```

### Interpolate at extra points from the sample points to determine prediction bias and precision
```{r pressure, echo=TRUE}                               
# Simple Kriging
simple.K <- krige(log10(avg_inf) ~ 1, frog_pos, grid_coord, model = m, beta = mean(log10(frog_pos@data$avg_inf)))
simple.k.d <- data.frame(simple.K)
                               
# Universal Kriging
universal.k <- krige(log10(avg_inf) ~ coords.x1+coords.x2, frog_pos, grid_coord, model = m)
universal.k.d <- data.frame(universal.k)
                               
# Ordinary Kriging
image(Ordinary.K["var1.pred"], col = cm.colors(20)) 
title("Ordinary Kriging Predictions") 
                               
# Simple Kriging 
image(simple.K["var1.pred"], col = cm.colors(20))
title("Simple Kriging Predictions") 
                               
# Universal Kriging
image(universal.k["var1.pred"], col = cm.colors(20))
title("Universal Kriging Predictions") 

# best semivariogram model  

k1 = autoKrige(log10(avg_inf) ~ 1, frog_pos, grid_coord) # Ordinary Kriging
plot(k1)
                               
v2 = autofitVariogram(log10(avg_inf) ~ 1, frog_pos_na)
plot(v2)
                               
m_co <- vgm(25307, "Gau", 3296055, 0.49) 
```

#Remove NAs
```{r}
frog_pos_na = frog_pos[which(!is.na(frog_pos@data$amphibia_1)),]
frog_pos_na = frog_pos_na[which(!is.na(frog_pos_na@data$amphibia_2)),]
frog_pos_na = frog_pos_na[which(!is.na(frog_pos_na@data$ph)),]
frog_pos_na = frog_pos_na[which(!is.na(frog_pos_na@data$watertemp)),]
frog_pos_na = frog_pos_na[which(!is.na(frog_pos_na@data$conductivi)),]
frog_pos_na = frog_pos_na[which(!is.na(frog_pos_na@data$totaldisso)),]
frog_pos_na = frog_pos_na[which(!is.na(frog_pos_na@data$salinity_p)),]
```

# Create semivariograms for co-variables SVL and weight

SVL
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "amphibian_svl", form = amphibia_1 ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "svl", model = m_co, fill.all=T)

k.c2 <- predict(g2, grid_coord, debug.level=0)
plot(variogram(g2), pl=T)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with SVL variable")
```

Weight
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "amphibian_weight", form = amphibia_2 ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "weight", model = m_co, fill.all=T)
                               
k.c2 <- predict(g2, grid_coord, debug.level=0)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with weight variable")
```

pH
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "PH", form = ph ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "PH", model = m_co, fill.all=T)
                               
k.c2 <- predict(g2, grid_coord, debug.level=0)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with Ph variable")
```

Water temperature
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "watertemp", form = watertemp ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "watertemp", model = m_co, fill.all=T)
                               
k.c2 <- predict(g2, grid_coord, debug.level=0)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with Watertemperature variable")
```                               

Water conductivity
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "conductivi", form = conductivi ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "conductivi", model = m_co, fill.all=T)
                               
k.c2 <- predict(g2, grid_coord, debug.level=0)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with conductivity variable")
```

Total dissolved solids
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "totaldisso", form = totaldisso ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "totaldisso", model = m_co, fill.all=T)
                               
k.c2 <- predict(g2, grid_coord, debug.level=0)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with Total dissolved solid variable")
```                               

Salinity
```{r pressure, echo=TRUE}
g2 <- gstat(NULL, id = "log10(avg_inf)", form = log10(avg_inf) ~ 1, data = frog_pos_na)
g2 <- gstat(g2, id = "salinity_p", form = salinity_p ~ 1, data = frog_pos_na)
                               
g2 <- gstat(g2, id = "salinity_p", model = m_co, fill.all=T)
                               
k.c2 <- predict(g2, grid_coord, debug.level=0)
                               
plot(variogram(g2), model=g2$model, pl=T)
                               
plot(k.c2, main = "Co-kriging with Salinity variable")
                               
```


